\sysfull

### 2018-04-16 paper/implementation.tex
Sections 5 and 6 revised

\subsection{Kernel Functions Implementation}

@@

\noindent \textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function. A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}. Upon reboot, the coalescing target is initialized according to the coalescing strategy. Then, the second phase of the commit is resumed, in case the last power failure occurred during this operation. The shadow buffer list is cleared, and the metadata of the volatile buffer is initialized. The pointer to the last coalescing task is recalled from non-volatile memory. Before running, \sys checks whether there is any partial execution to resume, in which case the volatile state, including the program counter, is restored. Otherwise, coalescing starts: the current budget is replenished; tasks are run, the history is increased and the current budget is decreased; when the latter is zero, coalescing stops, the pointer to the next task is saved and the commit phase starts. Coalescing resumes after commit phase two ends and the target is updated based on the coalescing strategy. \todo{Go over keywords (coalescing target, current budget, commit phases etc.) and make them consistent with the rest of the paper}{Amjad, Carlo}

\subsection{Paging Implementation}

\sys's memory virtualization is implemented using three buffers: (i) working (volatile) buffer, (ii) store (non-volatile) buffer and (iii) shadow (non-volatile) buffer. The store buffer holds the latest safe copy of all the pages, and it is read-only up to, and including, the ultimate commit. Upon reboot, the working buffer is empty, and it gets filled with pages from non-volatile memory on demand. All reads and writes to protected variables happen in volatile memory. The shadow buffer serves as an intermediate non-volatile buffer where dirty, volatile pages are copied during the intermediate commit. Moreover, it contains pages evicted from SRAM when a \emph{full page fault} occurs.

\textbf{Page Allocation and Page Tags.} When a \sys application accesses a protected variable with either \texttt{RP} or \texttt{WP}, \sys has to search for the variable in RAM first, and then, if not found, in non-volatile memory. In reality, the page manager searches for pages, not single variables. A page search can be very fast if each page is identified by a unique tag, and if every variable belonging to a page bears the tag in its address. Thus, retrieving a page tag from a protected variable's address boils down to extracting the upper bits of the address itself. The remaining lower bits denote the variable's offset in its page. The complete process is depicted in Figure~\ref{figure:coala_page_tags} \todo{this figure doesn't really show page tags and offsets, but paging in general}.
We note that, in order to obtain unique page tags, some alignment requirements have to be met. First, the page size $S$ has to be a power of two. Moreover, pages have to be aligned by $S$. Note that, by aligning pages by $2^n$, the alignment is also valid for all sizes $2^x$, where $x \leq n$.

@@

\textbf{Page Fault and Page Swap.} When a page is not found in RAM, a \emph{page fault} occurs, and it has to be fetched from FRAM. In case that page was in the working buffer and it had been evicted, its most recent non-volatile copy resides in the shadow buffer. For this reason, \sys searches this buffer first. If the requested page is not found in the shadow buffer, then it will be pulled from its store location. To make room for the requested page in SRAM, a volatile page might have to be evicted. If the victim page is dirty, we say that a \emph{full page fault} occurs, and the victim page has to be copied to the shadow buffer. Page moves are accelerated by the MCU's direct memory access (DMA) module.

>>>>

\subsection{Kernel Functions}

@@

\textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function. A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}. Upon reboot, the coalescing target is initialized according to the coalescing strategy. Then, the second phase of the commit is resumed, in case the last power failure occurred during this operation. The shadow buffer list is cleared, and the metadata of the volatile buffer is initialized. The pointer to the last coalescing task is recalled from non-volatile memory. Before running, \sys checks whether there is any partial execution to resume, in which case the volatile state, including the program counter, is restored. Otherwise, coalescing starts: the current budget is replenished; tasks are run, the history is increased and the current budget is decreased; when the latter is zero, coalescing stops, the pointer to the next task is saved and the commit phase starts. Coalescing resumes after commit phase two ends and the target is updated based on the coalescing strategy. \todo{Go over keywords (coalescing target, current budget, commit phases etc.) and make them consistent with the rest of the paper}{Amjad, Carlo}

\subsection{Paging}

As described in Section~\ref{sec:memory_virtulaization}, \sys's memory virtualization is implemented using three buffers: \texttt{working} (volatile) buffer, \texttt{\underline{store}} (non-volatile) buffer and \texttt{\underline{shadow}} (non-volatile) buffer. The \texttt{\underline{store}} buffer holds the latest safe copy of all the pages, and it is read-only up to, and including, the ultimate commit. Upon reboot, the \texttt{working} buffer is empty, and it gets filled with pages from non-volatile memory on demand. All reads and writes to protected variables happen in volatile memory. The \texttt{\underline{shadow}} buffer serves as an intermediate non-volatile buffer where dirty, volatile pages are copied during the intermediate commit. Moreover, it contains pages evicted from SRAM when a \emph{full page fault} occurs.

When a \sys application accesses a protected variable with either \texttt{RP} or \texttt{WP}, \sys has to search for the variable in RAM first, and then, if not found, in non-volatile memory. In reality, the page manager searches for pages, not single variables. A page search can be very fast if each page is identified by a unique tag, and if every variable belonging to a page bears the tag in its address. Thus, retrieving a page tag from a protected variable's address boils down to extracting the upper bits of the address itself. The remaining lower bits denote the variable's offset in its page. The complete process is depicted in Figure~\ref{figure:coala_page_tags} \todo{this figure doesn't really show page tags and offsets, but paging in general}.
We note that, in order to obtain unique page tags, some alignment requirements have to be met. First, the page size $S$ has to be a power of two. Moreover, pages have to be aligned by $S$. Note that, by aligning pages by $2^n$, the alignment is also valid for all sizes $2^x$, where $x \leq n$. For our experiments, we used 32-, 64-, 128- and 256-byte pages. 16\,KB were allocated in non-volatile memory for \texttt{\underline{shadow}} and \texttt{\underline{store}} (8\,KB each), and the volatile page buffer was 1\,KB, half the total SRAM size. The main advantage of having large pages is searches through page buffers are shorter. On the other hand, page swaps are more bulky.

@@
